---
title: "Asset Pricing 2 Project"
author: "Jia Ru"
date: "2016-4-5"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      results='hold',
                      cache = TRUE
                      )
```

##  1. Replicate exactly the same results as the examples for pricing barrier option in the PPTs.

We assume FX rate follows a geometric Brownian motion
$$ dS_t = \mu S_t dt + \sigma S_t dW_t $$
To simulate this path, we divide the time period to $N$ small intervals of length $\Delta t = T/N$, and discretize the SDE by
$$ S_{t+\Delta t} - S_t = S_t(\mu \Delta t + \sigma \sqrt{\Delta t}Z_t) $$
where $Z$ is standard normal variable.    
Follow the algorithm on Slides, I write the code:

```{r 1., eval=F}

rm(list = ls())

M <- 10000
N <- 1000
TT <- 1
delta_t <- TT/N
mu <- 0.03
sigma <- 0.12
U <- 1.71
L <- 1.29
X <- 1.4
S0 <- 1.5

f <- function(foo){
  S <- rep(NA,N)
  S[1] <- S0
  Z <- rnorm(N)
  for (t in 2:N){
    S[t] <- S[t-1] * (1 + mu*delta_t + sigma*sqrt(delta_t)*Z[t])
    if (N*1/12<t & t<N*6/12 & S[t]>U) return(0)
    if (N*8/12<t & t<N*11/12 & S[t]<L) return(0)
    if (t>N*11/12 & (S[t]>1.8|S[t]<1.3)) return(0)
  }
  #plot(x=1:N, y=S, type =  "l")
  payoff <- max(0,S[N]-X)
  return(payoff)
}

library(plyr)
library(doMC)
registerDoMC(8)

payoff_set <- 
  laply(.data = as.list(1:M), 
        .fun = f, 
        .parallel = TRUE,
        .progress = "text"
        )

p_mean <- mean(payoff_set)/(1+mu)
p_se <- sd(payoff_set)/sqrt(M)

p_mean
p_se

```

run the code I got the results: mean of option price is 0.08721852, Standard Error is 1.053E-3.

run the code 5 times, I got almost same results:
\begin{tabular}{|c|c|c|c|c|c|}
	\hline  & 1 & 2 & 3 & 4 & 5 \\ 
	\hline mean & 0.08809493 & 0.08584553 & 0.09035277 & 0.08712246 & 0.08908662 \\ 
	\hline SE & 0.001068521 & 0.001064478 & 0.001094953 & 0.001069485 & 0.001080291 \\ 
	\hline 
\end{tabular} 

## 2. Price European call option using Monte Carlo simulation, compare with result from analytic B-S formula. This procedure is called "benchmarking".

First, use Monte Carlo simulation.

```{r 2.MC}

rm(list = ls())

set.seed(12345)
S <- 1.5
K <- 1.4
sigma <- 0.12
r <- 0.03
ttm <- 1  # time to maturity

num.sim <- 10000

R <- (r-0.5*sigma^2)*ttm 
SD <- sigma*sqrt(ttm)
TTM.price<-S*exp(R+SD*rnorm(num.sim))
TTM.call<-pmax(0,TTM.price-K)
PV.call<-TTM.call*(exp(-r*ttm))

MC.call <- mean(PV.call)
MC.call

```
The Monte Carlo simualtion result of call opiton price is 0.1609742

Next, use B-S formula to calculate the European call option price
$$ C_{B-S} = S\Phi(d_{+})-Ke^{-r(T-t)}\Phi(d_{-}) $$

```{r 2.BS}

d1 <- (log(S/K)+(r+0.5*sigma^2)*ttm)/(sigma*sqrt(ttm)) # d+
d2 <- d1-sigma*sqrt(ttm) # d-

BS.call <- S*pnorm(d1)-K*exp(-r*ttm)*pnorm(d2)
BS.call

```

As the above output shows, the Black-Scholes call option value is 0.1610577. which is pretty close to the MC call option value of 0.1609742


## 3. Price this barrier option by Heston stochastic volatility using Monte Carlo simulation, also plot one sample path which hits the barrier and one path that never hits the barrier.

Heston's SV model is an extension of B-S model where the instantaneous volatility is itself a stochaastic process:
$$ \frac{dS_t}{S_t}=\mu dt +\sqrt{v_t}dW_t^1 $$
$$ dv_t = \kappa(\theta-v_t)dt + \sigma\sqrt{v_t}dW_t^2 $$
$$ dW_t^1 dW_t^2 = \rho dt $$
Discretize:
$$ \log S_{t+\Delta t} - \log S_t = (r-0.5v_t)\Delta t +\sqrt{v_t}\sqrt{\Delta t} Z_{t+1}^S $$
$$	\log v_{t+\Delta t} - \log v_t = \frac{1}{v_t}(\kappa(\theta-v_t)-\lambda v_t -0.5\sigma^2)\Delta t + \sigma \frac{1}{v_t}\sqrt{\Delta t} Z_{t+1}^v $$
$$  Z^v_t = \rho Z^s_t +\sqrt{1-\rho^2}Z_t 	$$

I write the following code to simulate the path on stock price and price the barriar call option.

```{r 3.HestonSV_f, eval=T}

rm(list = ls())

############  function that simulate ONE path  ################

f <- function(S0=100, K=90, r=0.025, ttm=1, v0=0.01, N=1000,
              rho=0.1, kap=2, theta=0.01, lambda=0.05, sigma=0.1,
              U=118, L=81, a=90, b=120)
{
  delta_t <- ttm/N
  
  Zs <- rnorm(N)  
  Zv <- rho*Zs + sqrt(1-rho^2)*rnorm(N)
  
  S <- c(S0,rep(NA,N-1)) # initalization of St
  lnS <- log(S)
  v <- c(v0,rep(NA,N-1)) # initalization of vt
  lnv <- log(v)
  for (t in 2:N) {
    # generate St one by one
    lnv[t] <- lnv[t-1] 
    + 1/v[t-1]*(kap*(theta-v[t-1])-lambda*v[t-1]-0.5*sigma^2)*delta_t 
    + sigma*1/sqrt(v[t-1])*sqrt(delta_t)*Zv[t]
    v[t] <- exp(lnv[t])
    lnS[t] <- lnS[t-1] + (r-0.5*v[t-1])*delta_t + sqrt(v[t-1])*sqrt(delta_t)*Zs[t]
    S[t] <- exp(lnS[t])
    
    # hit barriar ?
    if (
      (N*1/12<t && t<N*6/12 && S[t]>U) 
      || (N*8/12<t && t<N*11/12 && S[t]<L) 
      || (t>N*11/12 && (S[t]>b|S[t]<a)) 
    ){
      payoff <- 0
      dietime <- t
      break
      }
  }
    if (is.na(S[N])==FALSE) {
      payoff <- max(0,S[N]-K)
      dietime <- "didn't hit barriar"
      }
  
  return(list(St=S,dietime=dietime,payoff=payoff))
}
```

```{r 3.HestonSV_MC, eval=F}

############  Monte Carlo simulation  ################

set.seed(123456)
M <- 5000   # number of simulations

library(plyr)
library(doMC)
registerDoMC(8)  

payoff_set <- laply(
  .data = 1:M,
  .fun = function(foo) return(f()$payoff),
  .parallel = TRUE,
  .progress = "text"
)

rf <- 0.025  # f() function default risk free intrest rate is also 0.025
p_mean <- mean(payoff_set)/(1+rf) 
p_se <- sd(payoff_set/(1+rf))/sqrt(M)

p_mean
p_se

```

run the code above I got the simulated barriar call option price as 9.948896

plot some representable sample paths:

```{r 3.plot, eval=T}

############  plot  ################

library(ggplot2)
U=118
L=81
a=90
b=120
N=1000
# month : 1,6,8,11

p <- ggplot() + 
  geom_blank(aes(x=1:N,y=seq(from = 75, to = 125, length.out = N))) +
  xlab("time line : one year, 1000 steps") + 
  ylab("stock price") +
  ggtitle("3. Heston stochastic volatility") +
  geom_segment(aes(x=1/12*N,xend=6/12*N,y=U,yend=U)) +
  geom_segment(aes(x=8/12*N,xend=11/12*N,y=L,yend=L)) +
  geom_segment(aes(x=11/12*N,xend=N,y=a,yend=a)) +
  geom_segment(aes(x=11/12*N,xend=N,y=b,yend=b))


for (i in 1:5) {
  if (i==1) set.seed(123)
  if (i==2) set.seed(777)
  if (i==3) set.seed(666)
  if (i==4) set.seed(112) # hit barriar
  if (i==5) set.seed(779) # hit barriar
  ff <- f()
  assign(paste0("S",i),ff$St)
  assign(paste0("v",i),ff$dietime)
}

p + 
  geom_line(aes(x=1:N, y=S1),col="blue") +
  geom_line(aes(x=1:N, y=S2),col="blue") +
  geom_line(aes(x=1:N, y=S3),col="blue") +
  geom_line(aes(x=1:N, y=S4),col="red") +
  geom_segment(aes(x=v4,xend=v4,y=S4[v4],yend=75),col="red") +
  geom_line(aes(x=1:N, y=S5),col="red") +
  geom_segment(aes(x=v5,xend=v5,y=S5[v5],yend=75),col="red")
  
```
  

## 4. Price barrier option by Merton's Poisson jump model using Monte Carlo simulation, also plot one sample path which hits the barrier and one path that never hits the barrier.

Merton's Poisson jump model is just standard Geometric Brownian Motion plus Poisson jumps:
$$ \frac{dS_t}{S_t} = (r-\lambda\bar{k}) dt + \sigma dW_t + kdq_t $$
where $q_t$ is compound Poisson process with intensity $\lambda$,and
$$ \ln(1+k) \sim N(\gamma, \delta^2), \quad \bar{k} = e^{\gamma+\delta^2}-1 $$
The solution is
$$ S_t = S_0 e^{(r-\lambda\bar{k}-0.5\sigma^2)t + \sigma W_t} \prod_{i=0}^{n(t)}(1+k_i) $$
where $n(t)$ is Poisson process with rate $\lambda$
Take log we have
$$ \ln S_t - \ln S_0 = (r-\lambda\bar{k}-0.5\sigma^2)t + \sigma W_t + \sum_{i=1}^{n(t)}\ln(1+k_i) $$
discretize:
$$ \ln S_{t+\Delta t} - \ln S_t = 
(r-\lambda\bar{k}-0.5\sigma^2)\Delta t 
+ \sigma \sqrt{\Delta t} Z_t 
+ \Bigg[ \sum_{i=1}^{n(t+\Delta t)}\ln(1+k_i) - \sum_{i=1}^{n(t)}\ln(1+k_i) \Bigg]$$
Other terms are easy to calculate except the term in bracket. We can use a property of Poisson process: for any $s$, the increment $n(t+s)-n(t)$ follows poisson distribution with parameter $s \lambda$. So the term in bracket can be calculated with command  `sum(rnorm(n=rpois(lambda*delta_t), mean=gamma, sd=delta))`. Note that $\ln(1+k_i)$ is i.i.n.

```{r 4.jump_f, eval=T}

rm(list=ls())

# ####################################################### #  
#     function that simulate ONE path of Stock price      #
# ####################################################### #  

f <- function(
  S0 = 100, K = 90, r = 0.025, sigma = 0.15, TT = 1, N = 1000, 
  lambda = 5, gamma = 0, delta = 0.08,    # parameters of jump
  U = 130, L = 70, m1 = 1, m2 = 6, m3 = 8, m4 = 11 
){
  delta_t <- TT/N 
  k_bar <- exp(gamma + delta^2)-1

  lnS <- c(log(S0),rep(NA,N-1))   # initalization of ln(S_t)
  for (t in 2:N){
  lnS[t] <- lnS[t-1] + 
    (r  - lambda*k_bar - 0.5*sigma^2)*delta_t +
    sigma*sqrt(delta_t)*rnorm(1) + 
    sum(rnorm(rpois(n=1, lambda=lambda*delta_t), mean = gamma, sd = delta))  
  # hit barriar ?
  S <- exp(lnS)
  if (
    (S[t]>U && t/N>m1/12 && t/N<m2/12) 
    || (S[t]<L && t/N>m3/12 && t/N<m4/12) 
  ) {
    payoff <- 0; dietime <- t; break
  } else {
      payoff <- max(0,S[N]-K)
      dietime <- N
    } # end if
  } # end for
  
  return(list(S=S,dietime=dietime,payoff=payoff))
} 
```

```{r 4.jump_MC, eval=F}

# ####################################################### #  
#          Monte Carlo simulation
# ####################################################### #  
library(plyr)
library(doMC)
registerDoMC(8)

M <- 5000
N <- 1000
r <- 0.025

payoff_set <- laply(
  .data = 1:M,
  .fun = function(foo) return(f(r=r, N=N)$payoff),
  .parallel = T,
  .progress = "text"
)

p_mean <- mean(payoff_set)/(1+r) 
p_se <- sd(payoff_set/(1+r))/sqrt(M)

p_mean
p_se

```

run the code 5 times, I got almost same results:

\begin{tabular}{|c|c|c|c|c|c|}
	\hline  & 1 & 2 & 3 & 4 & 5 \\ 
	\hline mean & 11.6571 & 11.23386 & 11.85368 & 11.8501 & 11.7291 \\ 
	\hline SE & 0.2205481 & 0.2232357 & 0.2352804 & 0.2127964 & 0.2227964 \\ 
	\hline 
\end{tabular} 

Five plots of simulated path :
```{r 4.jump_plot, eval=T}
# ####################################################### #  
#          Visualization
# ####################################################### #  

library(ggplot2)

N=1000
m1=1; m2=6; m3=8; m4=11
U=130; L=70

p <- ggplot() + 
  geom_blank(aes(x=1:N,y=seq(from = 60, to = 135, length.out = N))) +
  xlab("time line : one year, 1000 steps") + 
  ylab("stock price") +
  ggtitle("4. Merton's Poisson jump") +
  geom_segment(aes(x=m1/12*N,xend=m2/12*N,y=U,yend=U)) +
  geom_segment(aes(x=m3/12*N,xend=m4/12*N,y=L,yend=L)) 


for (i in 1:5) {
  if (i==1) set.seed(1)
  if (i==2) set.seed(2)
  if (i==3) set.seed(777)
  if (i==4) set.seed(4) # hit barriar
  if (i==5) set.seed(5) # hit barriar
  ff <- f()
  assign(paste0("S",i),ff$S)
  assign(paste0("d",i),ff$dietime)
}

p + 
  geom_line(aes(x=1:N, y=S1),col="darkolivegreen3") +
  geom_line(aes(x=1:N, y=S2),col="darkolivegreen") +
  geom_line(aes(x=1:N, y=S3),col="darkolivegreen4") +
  geom_line(aes(x=1:N, y=S4),col="red") +
  geom_segment(aes(x=d4,xend=d4,y=S4[d4],yend=60),col="red") +
  geom_line(aes(x=1:N, y=S5),col="red") +
  geom_segment(aes(x=d5,xend=d5,y=S5[d5],yend=60),col="red2")

```












## 5. Price discrete-monitored barrier options (e.g. at the closing of the last day of each month or each day), and compare with the price under the continuous-time version.
Here I examine 5 cases: Benchmark (no barriar), monitored monthly, weekly, daily and continuously. By intuition, the (barriar) European call option price will decrease along these 5 cases. I will simulate the underlying stock price path and replicate it M=5000 times, draw the boxplot of simulated option price for each case.


```{r}

rm(list = ls())

# ####################################################### #  
#         f():  simulate ONE path of Stock price          #
# ####################################################### #  

f <- function(
  S0 = 100, K = 90, r = 0.025, sigma = 0.15, TT = 1, N = 12*30*24, 
  U = 120, L = 85, m1 = 1, m2 = 6, m3 = 8, m4 = 11,
  monitor = "benchmark"  # or monthly, weekly, daily)
){
  delta_t <- TT/N
  S <- c(S0,rep(NA,N-1))
  lnS <- log(S)
  Z <- rnorm(N)
  for (t in 2:N) {
    # 1. generate next S[t]
      lnS[t] <- lnS[t-1] + (r-0.5*sigma^2)*delta_t + sigma*sqrt(delta_t)*rnorm(1)
      S[t] <- exp(lnS[t])
    
    # 2. CHECK whether it's time to monitor, if not, `next`
    if (monitor=="benchmark") next
    Dcheck <- (monitor=="d" && t%%24==0)   # is it time to Daily monite?
    Wcheck <- (monitor=="w" && t%%(24*7)==0)   # weekly?   
    Mcheck <- (monitor=="d" && t%%(24*30)==0)   # Monthly? 
    if ( !(Dcheck|Wcheck|Mcheck) ) next
    
    # 3. GO TO monitor
    if (
      (N*m1/12<t && t<N*m3/12 && S[t]>U)
      ||(N*m2/12<t && t<N*m4/12 && S[t]<L)
      ) {
      payoff <- 0; dietime <- t; break
    } 
  } # End for
  if (t==N) {payoff <- max(0,S[t]-K); dietime="Not die"}
  
  return(list(S=S, dietime=dietime,payoff=payoff))
} # End function


# ####################################################### #  
#             simulate option price in 5 cases            #
# ####################################################### #  




```

